pollen_countR

### 2023-02-21 script/manuscript_MEE.Rmd
fully integrated reads!

--- 
title: Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees
author:  |
    |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
    |  $^4$ Australia post fire project, ----, ----, ----, AU  
    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
abstract:  |  
  |  
  1) DNA Barcoding has been successful for the rapid analysis of ecological assemblages. Although barcoding in the plant kingdom barcoding has been more difficult than others, and hence may begin to lag behind other kingdoms. 
  2) Here we test the utilization of Angiosperms 353 probes to barcode plant species found in pollen loads collected from Queen Bumble Bees. 
  3) To verify the accuracy for this barcoding system we compared the data to museum species, observation studies, and species distribution modelling to identify likely candidate species. 
  4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which may use increase the alignment algorithms minimizing need for large computational power, and run time. 
  5) We show that the Angiosperms 353 probes, which are currently being used in the largest ever plant systematic endeavor, offers significant promise to metagenomic approaches. 
  6) Understanding plants in ecological contexts and understandings of their synecology.
keywords: |
  DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species distribution model
output:
  word_document: default
  toc: no
  pdf_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: yes
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

@@

logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
  filter(occurrence == 0) 
# summary(logR_dat$n)
# sum(logR_dat$n)

@@

pmorph <- read.csv('../data/morpho_pollen.csv') 
blst_seqs <- distinct(blst, sample.id)

@@

rm(mertensia_samples, blst_aster, blst_aster1, blst_aster2, aster1_samples, 
   aster2_samples, palyn_cnt, blst_mertensia, aster_samples)
```

Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into. *APPENDIX 7*. From the 60 samples that were counted and based on rarefaction we **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*. The number of counted pollen grains in each sample range from (*MIN* - 16,293, $\bar{x}$ = 2788.685, Mdn = 1453).

@@

```{r}
rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
```

@@

@@

    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 4 ~ plant.species,                               ## COMPLETE ##

    sample.id == 6 ~ plant.species, # has a delphinium split
    # Symphyotrichum foliaceum check on day 173

    sample.id == 9 ~ plant.species, # has both Delphinium # Symphyotrichum foliaceum check on day 176

    sample.id == 10 ~ plant.species, # has many late season taxa

    sample.id == 14 ~ plant.species,  # # Symphyotrichum foliaceum check on day 165

    sample.id == 15 ~ plant.species, ## has many mysterious Rosids @ Gothic Spires  - call potentilla?

    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', # & Ericaceae
    sample.id == 18 ~ plant.species, # Erigeron grandiflorus on day 166 @ Frogs Pond
    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', # multiple Onagraceae
    sample.id == 19 ~ plant.species, # 
    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
    sample.id == 20 ~ plant.species, # Erigeron grandiflorus doy 165 @ Avery Wash
    sample.id == 21 ~ plant.species, # Erigeron grandiflorus & cirsium parryi doy 165 @ Gothic Spires

    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 23 ~ plant.species, # has both Viola             ## COMPLETE ##
    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,                              ## COMPLETE ##
    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,                              ## COMPELTE ##
    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,                              ## COMPLETE ##
    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 27 ~ plant.species, # has both Hydrophyllum      ## COMPLETE ##
    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,                              ## COMPLETE ##
    sample.id == 29 ~ plant.species,                              ## COMPLETE ##
    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 30 ~ plant.species,                              ## COMPLETE ##
    sample.id == 30 & Genus == 'Erigeron' ~ 'Erigeron elatior',
    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 32 ~ plant.species,                              ## COMPLETE ##
    sample.id == 33 ~ plant.species, # has both Viola
    # Symphyotrichum foliaceum check on day 166

    sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 35 ~ plant.species, # has both Viola             ## COMPLETE ##

    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  
    sample.id == 36 ~ plant.species,                              ## COMPLETE ##

    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 39 ~ plant.species, # has both hydrophyllum      ## COMPLETE ##
    sample.id == 40 ~ plant.species,                              ## COMPLETE ##
    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
    sample.id == 41 & Genus == 'Viola' ~ 'Viola adunca',          ## COMPLETE ##
    sample.id == 41 ~ plant.species,                              ## COMPLETE ##
    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,                             ## COMPLETE  ##
    sample.id == 44 ~ plant.species, # 44 has both delphinium

    sample.id == 45 ~ plant.species,         ## EARLY ASTER DAY 163 @ FROGS.POND ##
    sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 46 ~ plant.species,         ## EARLY ASTER DAY 163 @ FROGS.POND ##

    sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
    sample.id == 49 & Genus == 'Viola' ~ 'Viola.adunca',
    sample.id == 49 ~ plant.species,                              ## COMPLETE ##

    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 51 ~ plant.species,  ## Helianthella quinquenervis @ DOY 166 ##
    sample.id == 52 ~ plant.species,    ## Erigeron elatior  on DOY 163 @ AVERY WASH ##
    sample.id == 54 ~ plant.species,                             ## COMPLETE ##

    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 55 ~ plant.species,                              ## COMPLETE ##

    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
    sample.id == 56 ~ plant.species,          ## DOY 169 FROGS.POND Symphyotrichum foliaceum
                                              ## & Erigeron elatior ##
    sample.id == 60 & Genus == 'Viola' ~ 'Viola.adunca',          ## COMPLETE ##
    sample.id == 60 ~ plant.species,
    sample.id == 60 & Genus == 'Viola' ~ 'Viola',                 ## COMPLETE ##
    sample.id == 65 ~ plant.species,

@@

rm(ranunc_splash, blst_expert)

@@

@@

>>>>

--- 
title: Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees
author:  |
    |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
    |  $^4$ Australia post fire project, ----, ----, ----, AU  
    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
abstract:  |  
  |  
  1) DNA Barcoding has been successful for the rapid analysis of ecological assemblages. Although barcoding in the plant kingdom barcoding has been more difficult than others, and hence may begin to lag behind other kingdoms. 
  2) Here we test the utilization of Angiosperms 353 probes to barcode plant species found in pollen loads collected from Queen Bumble Bees. 
  3) To verify the accuracy for this barcoding system we compared the data to museum species, observation studies, and species distribution modelling to identify likely candidate species. 
  4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which may use increase the alignment algorithms minimizing need for large computational power, and run time. 
  5) We show that the Angiosperms 353 probes, which are currently being used in the largest ever plant systematic endeavor, offers significant promise to metagenomic approaches. 
  6) Understanding plants in ecological contexts and understandings of their synecology.
keywords: |
  DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species distribution model
output:
  pdf_document: default
  toc: no
  word_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: yes
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

@@

logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
  filter(occurrence == 0)

@@

pmorph <- read.csv('../data/morpho_pollen.csv') %>% 
  filter(!sample.id %in% c(45, 54))
blst_seqs <- distinct(blst, sample.id)

@@

rm(mertensia_samples, blst_aster, blst_aster1, blst_aster2, aster1_samples, 
   aster2_samples, palyn_cnt, blst_mertensia, aster_samples)

grains_counted <- pmorph %>% 
  group_by(sample.id) %>% 
  summarize(grains_counted = sum(sum)) %>% 
  pull(grains_counted)

Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into. *APPENDIX 7*. From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), **all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.** The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, Mdn = `r median(grains_counted)`).

@@

```{r}
rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected, grains_counted)
```

@@

# write.csv(blst, '../data/blst4experts.csv' , row.names = F)

@@

    sample.id == 3 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
    sample.id == 3 & Family %in% c('Rosaceae', 'Onagraceae') ~  'Potentilla.pulcherrima',
    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,       ## COMPLETE ##
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 4 ~ plant.species,                               ## COMPLETE ##
    sample.id == 6 ~ plant.species,                               ## COMPLETE ##
    sample.id == 9 & TAXON_NEW == 'Symphyotrichum foliaceum' ~ 'Erigeron speciosus',
    sample.id == 9 ~ plant.species,                               ## COMPLETE ##

    sample.id == 10 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
  #  sample.id == 10 & Family == 'Onagraceae' ~ ''
    sample.id == 10 ~ plant.species, # has many late season taxa

    sample.id == 14 & Genus == 'Symphyotrichum' ~ 'Senecio.integerrimus',
    sample.id == 14 ~ plant.species,                              ## COMPLETE ##
    sample.id == 15 & Family %in% c('Onagraceae', 'Rosaceae')  ~ 'Potential pulcherrima',
    sample.id == 15 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 15 ~ plant.species,                              ## COMPLETE ## 
    sample.id == 18 & Family == 'Ericaceae' ~ 'Ericaceae', # ONLY TO FAMILY
    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 18 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 18 ~ plant.species,                              ## COMPLETE ##

    sample.id == 19 & Family == 'Onagraceae' ~ 'Epilobium',
    sample.id == 19 & Genus == 'Pseudognaphalium' ~ 'Senecio.integerrimus',
    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri',
    sample.id == 19 ~ plant.species,                              ## COMPLETE ##
    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
    sample.id == 20 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 20 ~ plant.species,                              ## COMPLETE ##
    sample.id == 21 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 21 ~ plant.species,                              ## COMPLETE ##
    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 23 ~ plant.species,                              ## COMPLETE ##
    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,                              ## COMPLETE ##
    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,                              ## COMPELTE ##
    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,                              ## COMPLETE ##
    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 27 ~ plant.species,                              ## COMPLETE ##
    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,                              ## COMPLETE ##
    sample.id == 29 ~ plant.species,                              ## COMPLETE ##
    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 30 ~ plant.species,                              ## COMPLETE ##
    sample.id == 30 & Genus == 'Erigeron' ~ 'Erigeron elatior',
    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 32 ~ plant.species,                              ## COMPLETE ##

    sample.id == 33 ~ plant.species, # Symphyotrichum foliaceum check on day 166

    sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 35 ~ plant.species,                              ## COMPLETE ##
    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  
    sample.id == 36 ~ plant.species,                              ## COMPLETE ##
    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 39 ~ plant.species,                              ## COMPLETE ##
    sample.id == 40 ~ plant.species,                              ## COMPLETE ##
    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
    sample.id == 41 & Genus == 'Viola' ~ 'Viola adunca',          ## COMPLETE ##
    sample.id == 41 ~ plant.species,                              ## COMPLETE ##
    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,                              ## COMPLETE ##
    sample.id == 44 ~ plant.species,
    sample.id == 45 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 45 ~ plant.species,                              ## COMPLETE ##
    sample.id == 46 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 46 ~ plant.species,                              ## COMPLETE ##
    sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
    sample.id == 49 & Genus == 'Viola' ~ 'Viola.adunca',
    sample.id == 49 ~ plant.species,                              ## COMPLETE ##

    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 51 ~ plant.species,  ## Helianthella quinquenervis @ DOY 166 ##
    sample.id == 52 ~ plant.species,    ## Erigeron elatior  on DOY 163 @ AVERY WASH ##

    sample.id == 54 ~ plant.species,                              ## COMPLETE ##
    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 55 ~ plant.species,                              ## COMPLETE ##
    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
    sample.id == 56 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
    sample.id == 56 ~ plant.species,                              ## COMPLETE ##
    sample.id == 60 & Genus == 'Viola' ~ 'Viola.adunca',          ## COMPLETE ##
    sample.id == 60 ~ plant.species,
    sample.id == 65 & Genus == 'Viola' ~ 'Viola.praemorsa',       ## COMPLETE ##
    sample.id == 65 ~ plant.species,

@@

rm(ranunc_splash)

@@

# we will split the reads for the Cynoglossum detected in sample 3.

blst_reclass_expert <- blst_reclass_expert %>% 
  filter(sample.id == 3 & TAXON_EXPERT == 'Cynoglossum pringlei') %>% 
  slice(rep(1:n(), each=2)) %>% 
  mutate(Prcnt_seqs_Reclass = Prcnt_seqs_Reclass/2,
         dummyID = 1:n(),
         TAXON_EXPERT = if_else(dummyID == 1, 'Mertensia fusiformis', 'Mertensia cilita')) %>% 
  select(-dummyID) %>% 
  bind_rows(., filter(
              blst_reclass_expert, sample.id == 3 & TAXON_EXPERT != 'Cynoglossum pringlei'),
            filter(blst_reclass_expert, sample.id != 3))

@@

Regarding limitations of morphological data we suspect that there were two morphotypes of pollen identified as Ericaceae Juss. were actually Onagraceae (Samples 19 & 44), based on molecular results.

### 2023-02-21 script/pollen_countr3.Rmd
fully integrated reads!

pollen_df_new.1 <- pollen_df_new.1 %>%
  group_by(sample.id) %>% # form groups of each microscope slide. 
  mutate(count = as.numeric(count)) %>% # ensure that the values in this columns are represented as numbers to the computer
  mutate(count = replace_na(count, 0)) %>% # remember in our sample design NA's are 0's
  mutate(group_sum = sum(count)) %>% # see which slides have been counted...
  drop_na(morphotype) %>% 
  filter(group_sum >= 1) %>% # find only the slides which we have counted pollen grains in .
  mutate(
    morphotype = str_to_upper(morphotype),
    morphotype = str_replace(morphotype, "CAN\'T_DISTINGUISH", "UNKNOWN"),
    morphotype = str_replace(morphotype, "INDISTINGUISHABLE_CLUMP", "UNKNOWN"),
    morphotype = str_replace(morphotype, "NER", "UNKNOWN"),
    morphotype = str_replace(morphotype, "DEHYDRATED", "UNKNOWN"),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]5B[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, '5B', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'ASTERACEAE-2', 'ASTERACEAE_2'),
    morphotype = str_replace(morphotype, "SALIX", "D"),
    morphotype = str_replace(morphotype, 'C4', 'C'),
    morphotype = str_replace(morphotype, 'OENOTHERA', 'ONAGRACEAE'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]D[)]', 'D'),
    morphotype = str_replace(morphotype, 'OTHER_B', 'B'),
    morphotype = str_replace(morphotype, 'OTHER_C', 'C'),
    morphotype = str_replace(morphotype, "C\\_\\(LINUM\\)", "LINUM"),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]RANUNCULUS[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]11B[/]D[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]4_APERTURES[)]', 'UNKNOWN'),
    morphotype = str_replace_all(morphotype, " ", "_"),
    morphotype = str_remove(morphotype, "\\_\\(11B\\)"),
    morphotype = str_replace_all(morphotype, "UNKNWOWN", "UNKNOWN")) %>% 
  filter(!morphotype %in% c("NO_FJ", "LOW_FJ", "LITTLE_FJ", "LESS_FJ", 'LILTTLE_FJ')) %>%  # we cannot say we counted or scored these cells. 
  filter(!morphotype %in% c(4,8, 11)) %>% 
  mutate(plot.id = rleid(sample.id, transect, cell))

>>>>

pollen_df_new.1 <- pollen_df_new.1 %>%
  group_by(sample.id) %>% # form groups of each microscope slide. 
  mutate(count = as.numeric(count)) %>% # ensure that the values in this columns are represented as numbers to the computer
  mutate(count = replace_na(count, 0)) %>% # remember in our sample design NA's are 0's
  mutate(group_sum = sum(count)) %>% # see which slides have been counted...
  drop_na(morphotype) %>% 
  filter(group_sum >= 1) %>% # find only the slides which we have counted pollen grains in .
  mutate(
    morphotype = str_to_upper(morphotype),
    morphotype = str_replace(morphotype, "CAN\'T_DISTINGUISH", "UNKNOWN"),
    morphotype = str_replace(morphotype, "INDISTINGUISHABLE_CLUMP", "UNKNOWN"),
    morphotype = str_replace(morphotype, "NER", "UNKNOWN"),
    morphotype = str_replace(morphotype, "DEHYDRATED", "UNKNOWN"),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]5B[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, '5B', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'ASTERACEAE-2', 'ASTERACEAE_2'),
    morphotype = str_replace(morphotype, "SALIX", "D"),
    morphotype = str_replace(morphotype, 'C4', 'C'),
    morphotype = str_replace(morphotype, 'OENOTHERA', 'ONAGRACEAE'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]D[)]', 'D'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]GROSSULARIACEAE[)]', 'GROSSULARIACEAE'),
    morphotype = str_replace(morphotype, 'OTHER_B', 'B'),
    morphotype = str_replace(morphotype, 'OTHER_C', 'C'),
    morphotype = str_replace(morphotype, "C\\_\\(LINUM\\)", "LINUM"),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]AMSONIA[?)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]RANUNCULUS[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]11B[/]D[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]4_APERTURES[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, '^UNKNOWN.*$', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'UNKNOWN_[(]UNKNOWN[)]', 'UNKNOWN'),
    morphotype = str_replace(morphotype, 'PEDICULARISUNKNOWN', 'PEDICULARIS'),
    morphotype = str_replace_all(morphotype, " ", "_"),
    morphotype = str_remove(morphotype, "\\_\\(11B\\)"),
    morphotype = str_replace_all(morphotype, "UNKNWOWN", "UNKNOWN")) %>% 
  filter(!morphotype %in% c("NO_FJ", "LOW_FJ", "LITTLE_FJ", "LESS_FJ", 'LILTTLE_FJ')) %>%  # we cannot say we counted or scored these cells. 
  filter(!morphotype %in% c(4,8, 11)) %>% 
  mutate(plot.id = rleid(sample.id, transect, cell))
